MS&E 226: Project Part 1 Code

### 2024-12-02 MS&E 226_Project Part 1_Code.Rmd
code with ridge regression as model selected

View(state_means)
View(type_means)
View(region_means)
variables_to_plot <- names(state_means)[-1]  # Exclude Utility.State
plot_list <- list()
```

```{r Utility.State plots, fig.show='hide', results='hide'}
# Loop through each variable and create a plot
for (var in variables_to_plot) {
  p <- ggplot(state_means, aes_string(x = "Utility.State", y = var)) +
    geom_bar(stat = "identity", fill = "green") +
    coord_flip() +
    labs(title = paste("Mean", var, "by Utility State"),
         x = "Utility State",
         y = paste("Mean", var)) +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 6, angle = 0),
          axis.title.y = element_text(size = 10),
          axis.title.x = element_text(size = 10),
          plot.title = element_text(size = 12, face = "bold"))

>>>>

View(state_means)
View(type_means)
View(region_means)
```

```{r Utility.State plots, fig.show='hide', results='hide'}
variables_to_plot <- names(state_means)[-1]  # Exclude Utility.State
plot_list <- list()
# Loop through each variable and create a plot
for (var in variables_to_plot) {
  p <- ggplot(state_means, aes_string(x = "Utility.State", y = var)) +
    geom_bar(stat = "identity", fill = "green") +
    coord_flip() +
    labs(title = paste("Mean", var, "by Utility State"),
         x = "Utility State",
         y = paste("Mean", var)) +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 6, angle = 0),
          axis.title.y = element_text(size = 10),
          axis.title.x = element_text(size = 10),
          plot.title = element_text(size = 12, face = "bold"))

### 2024-12-02 MS&E 226_Project Part 2_Code.Rmd
code with ridge regression as model selected

@@

>>>>

### Regression

```{r Regression prediction with winning modeling strategy, Random Forest Model}

column_names <- c(
  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Sources.Purchased",
  "Sources.Other", "Uses.Retail", "Uses.Resale", "Uses.No.Charge", "Uses.Consumed", "Uses.Losses",
  "Retail.Residential.Customers", "Retail.Commercial.Customers", "Retail.Industrial.Customers",
  "Retail.Transportation.Customers", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type"
)

# Select the relevant columns in the test set
holdout_pred <- holdout[, column_names]
# Apply the same transformation to the test set
holdout_pred$Log_Revenue_Total <- log1p(holdout$Revenue.Total)

# Predict on the test set
rf_test_predictions <- predict(rf_model, newdata = holdout_pred)

# Check for NA values in predictions or Log_Revenue_Total
if (any(is.na(rf_test_predictions)) || any(is.na(holdout_pred$Log_Revenue_Total))) {
    stop("Predictions or Log_Revenue_Total contain NA values.")
}

# Calculate RMSE
rf_test_rmse <- sqrt(mean((rf_test_predictions - holdout_pred$Log_Revenue_Total)^2))
cat(paste("Random Forest Test RMSE:", rf_test_rmse, "\n"))

@@

#### (a)
```{r 2a}
# Extract the final Ridge model using the best lambda
final_ridge_model <- ridge_model$finalModel

# Extract coefficients from the final model
coefficients <- coef(final_ridge_model, s = final_lambda_ridge)

# Print the coefficients
print(coefficients)

#### (b)
```{r}
# Select a subset of covariates
subset_column_names <- c(  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type")

# Create a new data frame with the subset of columns
df_subset <- df[, subset_column_names]

df_subset$Log_Revenue_Total <- log1p(df_subset$Revenue.Total)

# Train the Ridge Regression model using the subset of covariates
ridge_subset_model <- train(Log_Revenue_Total ~ ., 
                            data = df_subset, 
                            method = "glmnet",  
                            trControl = train_control,
                            metric = "RMSE", 
                            preProcess = "scale",  
                            tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 1, length = 100)))

```
```{r}
# Get the coefficients for the best model at the selected lambda
best_lambda <- ridge_model$bestTune$lambda
best_model_coeffs <- coef(ridge_model$finalModel, s = best_lambda)

# Print all coefficients for the best Ridge Regression model
cat("All coefficients for the best Ridge Regression model:\n")
print(best_model_coeffs)

# Get coefficients for the subset model
best_lambda_subset <- ridge_subset_model$bestTune$lambda
best_model_coeffs_subset <- coef(ridge_subset_model$finalModel, s = best_lambda_subset)

# Print all coefficients for the subset model
cat("All coefficients for the Ridge Regression model with a subset of covariates:\n")
print(best_model_coeffs_subset)
```
